Analysis of HTTPS Adoption and Third-party Tracking: \\A Swedish Perspective

### 2015-05-07 paper/paper.tex
Paper as submitted to IMC15

%%\documentclass{sig-alternate}
\documentclass{sig-alternate-10pt}
%%\documentclass{sig-alternate-2013}
%\newfont{\mycrnotice}{ptmr8t at 7pt}
%\newfont{\myconfname}{ptmri8t at 7pt}
%\let\crnotice\mycrnotice%
%\let\confname\myconfname%

@@

@@

\numberofauthors{2} %  in this sample file, there are a *total*
\author{
\alignauthor Joel Purra\\
       \affaddr{Link{\"o}ping University}\\
       \affaddr{Link{\"o}ping, Sweden}\\
       \email{mig@joelpurra.se}
\and
\alignauthor Niklas Carlsson\\
       \affaddr{Link{\"o}ping University}\\
       \affaddr{Link{\"o}ping, Sweden}\\
       \email{niklas.carlsson@liu.se}
}

@@

Both active third-party tracking (e.g., using scripts and plugins embedded within visited websites) and passive tracking 
(e.g., through network monitoring) are extensively used to gather knowledge about users internet activities and interests.  
While HTTPS can be used to protect against some privacy concerns raised by passive monitoring, prior work have not carefully 
characterized the HTTPS adoption and third-party tracking when using HTTP and HTTPS, respectively.  This paper presents a 
measurement analysis of the current HTTPS adoption and third-party tracking usage.  First, we develop and release the 
software of a novel measurement framework for automated, repeatable retrieval, and analysis of websites.  Second, 
we characterize the HTTPS adoption, HTTPS redirection dynamics, and use of secure/non-secure third-party resources 
from both a national (Swedish) and global perspective.  In total, we analyze more than 150,000 websites.  Third, 
we characterize the third-party tracking usage and the companies that may get access to user information when using 
HTTP and HTTPS.  Overall, we have found that users may be given a false sense of privacy when using HTTPS.  
They are often directed to non-secure subdomains, and even when served using HTTPS, the use of third-party 
(external) resources is high among all classes of domains, regardless if HTTP or HTTPS is used.  In fact, 
third-party usage is most prevalent among the most popular website, visited most often.  While Google has 
by far the greatest third-party coverage (more than 90\% for most website categories), we have found that 
there are many other domains serving third-party content that may fly under the radar.

%Both active third-party tracking (e.g., using scripts and plugins embedded within visited websites) and passive tracking (e.g., through network monitoring) are extensively used to collect information and knowledge about users internet activities and interests.  While HTTPS can be used to protect against privacy concerns raised by passive monitoring, prior work have not carefully studied the adoption of HTTPS and how much privacy protection against third-party tracking can be gained from using HTTPS.  This paper presents a measurement analysis of the current HTTPS adoption and third-party tracking usage.  First, we develop and release the software of a novel measurement framework for automated, repeatable retrieval, and analysis of websites and their third-party usage and HTTPS dynamics.  Second, we characterize the HTTPS adoption and use of secure/non-secure third-party resources from both a national (Swedish) and global perspective.  In total, we repeatedly analyze more than 150,000 websites.  Third, we characterize the third-party tracking usage and the companies that may get access to user information when using HTTP and HTTPS.  Overall, we have found that users may be given a false sense of privacy when using HTTPS.  They are often directed to non-secure subdomains, and even when served using HTTPS, the use of third-party (external) resources is high among all classes of domains, regardless if HTTP or HTTPS is used.  This is magnified by third-party usage being most prevalent among the most popular website, which people visit most often.  While Google has by far the greatest third-party coverage (more than 90\% for most website categories) among known tracker companies, we have found that there are many other domains that also server third-party content which may fly under the radar.

@@

% A category with the (minimum) three required fields
%\category{C.4}{Information Systems Organization}{Performance of Systems}
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
\category{C.2.2}{Information Systems Organization}{Network Protocols}[Applications] 
%\category{H.5.1}{Multimedia Information Systems}{Video}

@@

\keywords{Blah blah}
%%\keywords{Nonlinear streaming; Multipath streaming; HAS}

@@

We are living in an information society in which organizations constantly track our movements 
on the world-wide web, and use the collected (and stored) information to analyze and gain 
knowledge about us and our interests.  Website owners may use such knowledge to present personalized 
content that may better match user interests, advertisement firms may use the knowledge to sell 
targeted ads based on user interests, media analytics firms may use the data to verify 
advertisement-related statistics, and data brokers may package and sell the user data 
inferred from the user interests.

Web usage is both actively and passively tracked.  With active tracking, third-party 
scripts and plugins embedded within the visited websites, for example, are typically 
used to extract and collect information about a user's every click, the time spent on 
each page (even if inactive), the user environment (including operating system, window size, 
screen resolution, and color depth), mouse movements, scrollbar location, installed fonts, 
plugins, extensions, and any potential error reports, for example.  With passive measurements, 
network traffic can be monitored at any point in the network, allowing user actions and user 
interests to be inferred from the traffic patterns, HTTP-related information (such as URLs of 
visited websites, previously visited websites, the client's browser, operating system, and cookies, for example), 
and any other application specific information (including information sent to third-party trackers).

In parallel with a trend towards increased user tracking, there has also been an increase in the 
number of services that use HTTPS.  Such secure end-to-end connections (between the client and server) 
help limit the amount of information that can be gained from passive data collection along the network path, 
but does not protect against active third-party tracking enabled through embedded tracking code.  
When considering the current state-of-the art, and these trends in particular, 
a number of important question arise:  What is the status of the current HTTPS adoption?  
How much third-party tracking is deployed?  Does third-party tracking usage differ for early 
adapters of HTTPS? And, perhaps most importantly, how much privacy and protection against 
third-party tracking can be gained from conscious use of HTTPS?  As of today, there have 
only been some preliminary works touching upon the first two of these questions~\cite{BuMS11,NFL+14}, 
but no work has carefully addressed all four questions.

To carefully address these important and un-answered questions, this paper presents 
a novel measurement methodology and analysis of the current HTTPS and third-party tracking usage.  
Leveraging Swedish domain knowledge, the study presents results from both a national (Swedish) 
perspective as well as based on a global perspective (using globally popular service, for example).  
The paper makes three main contributions:

First, we develop and release the software of a novel measurement framework for automated, 
repeatable retrieval, and analysis of websites and their third-party usage and 
HTTPS dynamics.~\footnote{Both the open source project and datasets used in the paper 
are made public with this paper.   Code and datasets and can be found here: \url{http://joelpurra.com/projects/masters-thesis/}.}  
The software is built on open source software, uses a non-modified ``standard'' headless browser, 
and stores results into standardized HAR format.  To capture HTTPS adoption and redirections, 
domains are visited using both \texttt{http} and \texttt{https}, with and without the \texttt{www} prefix. 
Third-party resource usage is captured through post processing (including comparison with the tracker 
lists of the popular privacy tool \texttt{disconnect.me}\footnote{ADD FOOTNOTE LINK TO DICONNECT.ME}) 
and public suffix lists are used to identify domains and subdomains.

Second, we characterize the HTTPS adoption and use of secure/non-secure third-party resources from 
both a Swedish and global perspective.  We find that the HTTPS adoption among globally popular 
websites (10-30\%, 50\% for the very top) and lists of both popular and important Swedish 
websites (15-50\%) is much higher than for random domains (less than 1\%). Although it appears 
that popular websites are much further along in their adoption, the low overall values suggest 
that most websites are susceptible to passive eavesdropping anywhere along the network path.  
To make things worse for the privacy conscious user, we have found that many websites redirect 
their users from their secure (\texttt{https}) domain to their insecure (\texttt{http}) domain, 
and most secure \texttt{https} domains relies heavily on both internal and external (third party) 
resources delivered over HTTP (rather than HTTPS).  The use of third-party resources over HTTP 
shows that a secure, encrypted connection to a website often does not ensure privacy neither 
against active and passive tracking.  This may be concerning as users falsely might be lead to 
believe that such “secure connection” (by many browsers labeled using a padlock) provides 
protection against external tracking.

Third, we characterize the third-party tracking usage and the companies that may get access 
to user information when using HTTP and HTTPS.  Overall, the use of third-party (external) 
resources is high among all classes of domains, regardless if HTTP or HTTPS is used.  
For example, less than 7\% of the globally most popular domains use only internal resources.  
These numbers are even lower for the most popular Swedish websites within most websites categories, 
but greater among random domains (10-30\%), suggesting that users activates often can be tracked 
through third-party resource usage.  In fact, most websites also have at least one {\em known} 
tracker present: 53-72\% of random domains, 88-98\% of top websites, and 78-100\% of websites 
in the different Swedish top-categories.  While Google has by far the greatest third-party 
coverage among known tracker companies (e.g., Google has trackers on more than 90\% of the 
websites within the majority of the website categories considered here), we have found that 
there are many other third-party domains that may fly under the radar.  For example, 
Disconnect's blocking list only detects 10\% of external primary domains.  With most 
of these non-blocked third-party entries being third-party content providers, which are 
known to keep track of users across services, we expect that there will be a continued 
battle for the knowledge about our web activities.

The remainder of the paper is organized as follows. Section~\ref{sec:background} gives 
a brief background on online tracking.  Our methodology and measurement framework are 
described in Section~\ref{sec:method}.  Section~\ref{sec:https} presents our 
characterization results for HTTPS adoption and client redirections between 
secure and non-secure resources.  Section~\ref{sec:tracking} characterizes the 
third-party tracking landscape and how it differs between secure and non-secure 
domains. Finally, Section~\ref{sec:related} discusses related works, before 
Section~\ref{sec:conclusions} concludes the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}\label{sec:background}

@@

\begin{table*}[t]
\centering
\caption{Summary of domain lists.}
\label{tab:domains}
{\small
\begin{tabular}{|l|c|r|l|r|r|}\hline
\hline %% closing line
Domain list       & Date       & Total size  & Selection & Selection size & Unique\\\hline
.SE health status & 2014-03-27 & 980         & curated   & 915            & 915   \\\hline
.se zone          & 2014-07-10 & 1,318,000   & random    & 100,000        & 100,000 \\\hline
.dk zone          & 2014-07-23 & 1,260,000   & random    & 10,000         & 10,000 \\\hline
.com zone         & 2014-08-27 & 114,178,000 & random    & 10,000         & 10,000 \\\hline
.net zone         & 2014-08-27 & 15,096,000  & random    & 10,000         & 10,000 \\\hline
reach50           & 2014-09-01 & 50          & top       & 50             & 50 \\\hline
Alexa Top 1M      & 2014-09-01 & 1,000,000   & top       & 10,000         & 9,986 \\\hline
                  &            &             & random    & 10,000         & 9,959 \\\hline
                  &            &             & all .se   & 3,364          & 3 364 \\\hline
                  &            &             & all .dk   & 2,637          & 2,637 \\\hline
Total             &            & 132,852,050 &           & 156,907        & 156,045 \\\hline
\end{tabular}}
\vspace{-0pt}
\end{table*}

@@

%In this work, 
We perform a large-scale measurement campain in which 
the front page of a large set of domains are visited
using both \texttt{http} and \texttt{https}.
(We also considered both domain variations 
with and without the \texttt{www} prefix.)
%%Based on a list of domains, 
%%At a high level,
With the tool developed in this project,
%%\footnote{The browser does not have a visible window, as it is built for automation.}
%%browser \texttt{phantomjs} downloads
the front page of each domain is downloaded and parsed 
using a headless (without display window) \texttt{phantomjs} browser 
the same way a user's browser would.\footnote{
As part of the validation process,
the tool creates an image of the full website after the download is complete,
but does not display the results to any user.}
The URL of each requested resource is extracted, and associated with
the domain it was loaded from. Each object and domain is classified 
and prepared for post processing and analysis.

{\bf Swedish perspective:}  
First, our meaurements are performed from computers located in Sweden.
Second, we leverage the domain lists from the .SE Health Status report\cite{},
identified by the Internet Infrastructure Foundation (.SE) as the most important domains
to Swedish internet usage and operation.
Combined, these lists include approximately 1,000 domains in the
categories: counties, domain registrars, financial services, government-owned corporations
(GOCS), higher education, ISPs, media, municipalities, and public authorities.
Third, we consider the 50 global websites most visited by Swedish Internet users (reach50),
as well as randomly selected websites within the .se domain zone.

{\bf Global baseline:}  To set our findings in perspective and to broader our
conclusions we also consider globally popular websites (according to Alexa\footnote{
Alexa (official website), \url{www.alexa.com}.}) and randomly selected websites 
from the .com, .net, and .dk (Denmark) domain zones.
Table~\ref{tab:domains} summarizes the domain lists used.
%%and any bias due to geographic personalization by content providers and
%%Content Delivery Networks (CDNs), for example.
%%Client in Sweden and section of websites ...

{\bf Data collection and parallelization:}
HTTP/HTTPS traffic metadata such
as requested URLs and their HTTP request/response headers have been recorded in the HTTP
Archive (HAR) data format.
During our data collection,
multiple domains have been retrieved in parallel, 
with parallelism adjusted to fit each computers' capacity. 
To reduce the risk of intermittent errors,
%% – either in software, on the network or in the remote system – 
each failed access has been retried up to two times.

@@

{\bf Data extraction:}
A custom-built tool based on the command line JSON processor \texttt{jq}
is used to covert the HAR data into JSON-based HAR data with information about both requests and responses.
The extracted data includes protocol and hosts (e.g., from URLs), 
HTTP status, mime-type, referer and redirect values – both
for the origin domain's front page and any resources requests by it.

{\bf Domain identification and classification:}
We classify each requested resource along many dimensions, 
including its mime type, if it is retrieved securely,
and if it is downloaded from a sub-domain or from a known third-party tracker, for example. 
To determine the primary domain and subdomain(s) we use the public suffix list\footnote{REFERENCE/POINTER???} 
(used by browser vendors to implement domain-dependent security measures)
and match domains against it.  This is important to capture the relationships
of second-level domains not open for commercial registration (e.g., \texttt{.com.br}
under \texttt{.br}).  
We refer to resources requested from the main page
that are retrieved from the same domain, a subdomain
or a superdomain as internal requests.
Any other request is deemed as an external request.

{\bf Tracker identification:}
While there are public lists of known trackers, 
%%used by browser privacy tools, they 
these lists are often incomplete and out-of-date~\cite{ref35, ref22}.
In fact,  any external resource can be used for third-party tracking.
For example, even for static (non-script, non-executable) resources with no capabilities to
dynamically survey the user’s browser or OS can track users across domains using
the referer  HTTP header and custimized URI's, for example.
These lists are therefor perhaps best seen as lists of confirmed, recognized third-party trackers.

In our experiments, we do not block are requests.  Instead, URL are matched against known 
trackers during post processing.  For this analysis, we use the tracker list
used by the privacy tool Disconnect.me.  This list contains 2,149 domains,
each belonging to one of 980 organizations and five tracker categories.

>>>>

%%\documentclass{sig-alternate}
\documentclass[letterpaper]{sig-alternate-10pt}
%%\documentclass{sig-alternate-2013}
%\newfont{\mycrnotice}{ptmr8t at 7pt}
%\newfont{\myconfname}{ptmri8t at 7pt}
%\let\crnotice\mycrnotice%
%\let\confname\myconfname%

@@

\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}

@@

\numberofauthors{2} %  in this sample file, there are a *total*
\author{
\alignauthor Joel Purra\\
       \affaddr{Link{\"o}ping University, Sweden}\\
%%       \affaddr{Link{\"o}ping, Sweden}\\
       \email{mig@joelpurra.se}
\and
\alignauthor Niklas Carlsson\\
       \affaddr{Link{\"o}ping University, Sweden}\\
%%       \affaddr{Link{\"o}ping, Sweden}\\
       \email{niklas.carlsson@liu.se}
}

@@

%Both active third-party tracking (e.g., using scripts and plugins
%embedded within visited websites) and passive tracking
%%(e.g., through network monitoring)
Today, extensive use of third-party tracking services and passive 
%%tracking through 
traffic monitoring are both used to 
%%are extensively used to 
gather knowledge about users' internet activities and interests.  
With HTTPS providing some protection against passive monitoring,
privacy concerned users are increasingly selecting services 
%that are 
provided over HTTPS.
This paper presents a measurement-driven 
analysis of the current HTTPS adoption and its relationship
to third-party tracking usage.
%%addressing the lack of careful characterizing of .
%%While HTTPS can be used to protect against some privacy concerns raised by passive monitoring, 
%prior work have not carefully characterized the HTTPS adoption 
%and third-party tracking when using HTTP and HTTPS, respectively.  This paper presents a 
%measurement analysis of the current HTTPS adoption and third-party tracking usage.  
First, we develop and release the 
software of a novel measurement framework for automated, repeatable retrieval, and analysis of websites.  
Second, we characterize the HTTPS adoption, HTTPS redirection dynamics, 
and use of secure/non-secure third-party resources from both a national (Swedish) and global perspective.  
%Across domain classes, both nationally and internationally,
%we observe low HTTPS adoption, 
%many domains that redirect to insecure resources,
%and many domains that use a mix of secure and insecure connections.
%%In total, we analyze more than 150,000 websites.  
Third, we compare and contrast the third-party tracking usage,
including the tracker service and the companies that may get access to user information,
when using HTTP and HTTPS, respectively.
%%and characterization the tracker service and the companies that may get access to user information.
%% when using HTTP and HTTPS.  
Across both national and international domain categories
we observe low-to-modest (below 50\%) HTTPS adoption,
%We also observed 
many domains that redirect to insecure domains,
%that 
or
%%and many domains 
that use a mix of secure and insecure connections 
to download the resources making up the presented website.
%or that do both.
%While Google has by far the greatest third-party coverage (more than 90\% for most website categories), 
%we have found that there are many other third-party domains that that may fly under the radar.
%domains serving third-party content that may fly under the radar.
%%Overall, we have found that users may be given a false sense of privacy when using HTTPS.  
%%They are often directed to non-secure subdomains, and even when served using HTTPS, 
The use of third-party resources is high among all considered domain categories, 
regardless if HTTPS is used or not,
suggesting that users may be given a false sense of privacy when using HTTPS.
%%In fact, both HTTPS and third-party usage is most prevalent among the most popular website, visited most often.  
%%While Google has by far the greatest third-party coverage (more than 90\% for most website categories), we have found that 
%%there are many other domains serving third-party content that may fly under the radar.
While Google has by far the greatest third-party coverage (more than 90\% for most website categories),
we have found that there are many other third-party domains that that may fly under the radar.

%Both active third-party tracking (e.g., using scripts and plugins embedded within visited websites) and passive 
%tracking (e.g., through network monitoring) are extensively used to collect information and knowledge about users 
%internet activities and interests.  While HTTPS can be used to protect against privacy concerns raised by passive 
%monitoring, prior work have not carefully studied the adoption of HTTPS and how much privacy protection against 
%third-party tracking can be gained from using HTTPS.  This paper presents a measurement analysis of the current 
%HTTPS adoption and third-party tracking usage.  First, we develop and release the software of a novel measurement 
%framework for automated, repeatable retrieval, and analysis of websites and their third-party usage and HTTPS 
%dynamics.  Second, we characterize the HTTPS adoption and use of secure/non-secure third-party resources from 
%both a national (Swedish) and global perspective.  In total, we repeatedly analyze more than 150,000 websites.  
%Third, we characterize the third-party tracking usage and the companies that may get access to user information 
%when using HTTP and HTTPS.  Overall, we have found that users may be given a false sense of privacy when using 
%HTTPS.  They are often directed to non-secure subdomains, and even when served using HTTPS, the use of third-party 
%(external) resources is high among all classes of domains, regardless if HTTP or HTTPS is used.  This is magnified 
%by third-party usage being most prevalent among the most popular website, which people visit most often.  While 
%Google has by far the greatest third-party coverage (more than 90\% for most website categories) among known 
%tracker companies, we have found that there are many other domains that also server third-party content which 
%may fly under the radar.

@@

%% A category with the (minimum) three required fields
%%\category{C.4}{Information Systems Organization}{Performance of Systems}
%%\category{H.4}{Information Systems Applications}{Miscellaneous}
%%A category including the fourth, optional field follows...
%%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
%\category{C.2.2}{Information Systems Organization}{Network Protocols}[Applications] 
%%\category{H.5.1}{Multimedia Information Systems}{Video}

@@

%\keywords{Blah blah}
%%\keywords{Nonlinear streaming; Multipath streaming; HAS}

@@

We are living in an information society in which organizations constantly track our movements 
on the web, and use the collected (and stored) information to analyze and gain 
knowledge about us and our interests.  Website owners may use such knowledge to present personalized 
content that may better match user interests, advertisement firms may use the knowledge to sell 
targeted ads based on user interests~\cite{GEC+13,SLB+14,YZL+13,LSW+13}, media analytics firms may use the data to verify 
advertisement-related statistics, and data brokers may package and sell the user data 
inferred from the user interests~\cite{Armo14}.

Web usage is both actively and passively tracked.  With active tracking, 
third-party scripts and plugins embedded within the visited websites
%%for example,
are typically used to extract and collect information about a user's every click, 
the time spent on each page (even if inactive), and information
that can be used to uniquely identify each browser and client
(including operating system, window size, screen resolution, installed fonts, etc.).
%%the user environment (including operating system, window size, 
%%screen resolution, and color depth), mouse movements, scrollbar location, installed fonts, 
%%plugins, extensions, and any potential error reports, for example.  
While some trackers allow advanced fingerprinting through collection of client-side information,
much tracking is achieved through server-side tracking of the download of
relatively simple third-party contents such as CSS files, customized fonts, 
or javascript libraries that does not include tracking code themselves~\cite{AEE+14,MPS+13,MaMi12,RoKW12,NIK+12}.  
%All external resources are considered trackers, even if code isn't executed on the client. 
%CSS and fonts for example are major trackers for Google, as well as their hosting of javascript 
%libraries (such as jquery) which doesn't actually contain tracking code -- it's all passive, 
%server-side tracking and often served over HTTPS.
User activities can also be 
%passively 
tracked through 
passive traffic monitoring within the network or at a cloud-operated data center hosting some of the content.
%%can also be performed by passive monitoring of the end-to-end network traffic. 
%%network traffic can be monitored at any point in the network or at a cloud-operated data center, for example,
Here, unencrypted HTTP data such as header information 
%(e.g., URLs, previously visited websites, and the client's browser, OS, and cookies)
as well as application specific information (including information sent to third-party trackers)
could be used to identify users and track their web sessions.
%%allowing user actions and user interests to be inferred from the traffic patterns, HTTP-related information (such as URLs of 
%%visited websites, previously visited websites, the client's browser, operating system, and cookies, for example), 
%%and any other application specific information (including information sent to third-party trackers).

In parallel with a trend towards increased user tracking, 
there has also been an increase in the number of services that use HTTPS.  
HTTPS helps secure the end-to-end communication between the client and server
%%help 
and limits the information that can be passively collected along the network path. 
However, it does not protect against tracking enabled through embedded 
third-party content or tracking codes.  
When considering the current state-of-the art, and these trends in particular, 
a number of important question arise:  What is the status of the current HTTPS adoption?  
How much third-party tracking is deployed?  Does third-party tracking usage differ for 
%%early 
domains adopting HTTPS?  And, perhaps most importantly, how much privacy and protection against 
third-party tracking can be gained from conscious use of HTTPS?  
As of today, there have only been some preliminary works touching upon the first 
two of these questions~\cite{BuMS11,NFL+14}, 
but no work has carefully addressed all four questions.

To help address these important and un-answered questions, this paper presents 
a novel measurement methodology and analysis of the current HTTPS and third-party tracking usage.  
Leveraging Swedish domain knowledge, the study presents results from both a national (Swedish) 
and global perspective.
%% (using globally popular service, for example).  
The paper makes three main contributions.

First, we develop and release the software of a 
%novel
measurement framework for automated, 
repeatable retrieval, and analysis of websites and their third-party usage and 
HTTPS dynamics.\footnote{Both the open source project and datasets used in the paper 
are made public with this paper,
and can be found here:
%Code and datasets can be found here: 
\url{http://joelpurra.com/projects/masters-thesis/}.}  
The software is built on open source software, uses a non-modified ``standard'' headless browser, 
and stores results into standardized HAR format.  To capture HTTPS adoption and redirections, 
domains are visited using both \texttt{http} and \texttt{https}, with and without the \texttt{www} prefix. 
Third-party resource usage is captured through post processing (including comparison with the tracker 
lists of the popular privacy tool \texttt{disconnect.me}\footnote{Disconnect, \url{https://disconnect.me/}, May 2015.}) 
and public suffix lists are used to identify domains and subdomains.

Second, we characterize the HTTPS adoption and use of secure/non-secure third-party resources from 
both a Swedish and global perspective.  
%%We find that the HTTPS adoption among globally popular 
%%websites (10-30\%) and lists of both popular and important Swedish 
%%websites (15-50\%) is much higher than for random domains (less than 1\%). 
Although it appears that the globally most popular websites (10-30\% HTTPS adoption)
and popular and important Swedish websites (15-50\%) are much further along in their adoption
than random domains (less than 1\% adoption), the low overall values suggest 
that most websites are susceptible to passive eavesdropping anywhere along the network path.  
To make things worse for the privacy conscious user, we have found that many websites redirect 
their users from their secure (\texttt{https}) domain to their insecure (\texttt{http}) domain, 
and many of the domains that allow use of HTTPS still 
%% secure \texttt{https} domains 
relies heavily on some (internal and third-party)
resources used for the website to be delivered insecurely.
% over HTTP,rather than over HTTPS.
%%both internal and external (third-party) 
%resources delivered over HTTP (rather than HTTPS).  
%The use of third-party resources over HTTP shows that a secure, 
%encrypted connection to a website often does not ensure privacy neither 
%against active and passive tracking.  
This is concerning as users falsely might be lead to 
believe that such “secure connection” 
%(by many browsers labeled using a padlock) 
provides protection against external tracking,
while the client may still be tracked both through third-party resource usage
and by monitoring non-secure connections. 
%still allow both active-third party tracking
%and communication over non-secure connections.
%%both passive monitoring of non-secure connections and active-third party tracking.

Third, we characterize the third-party tracking usage and the companies that may get access 
to user information when trying to download the front page of each domain 
using HTTP and HTTPS.  Overall, the use of third-party (external) 
resources is high among all domain categories, regardless if HTTP or HTTPS is used.  
For example, less than 7\% of the globally most popular domains use only internal resources.  
These numbers are even lower for the most popular Swedish websites within most websites categories, 
%but greater among random domains (10-30\%), 
suggesting that users' activities often can be tracked 
through third-party resource usage.  In fact, most websites also have at least one {\em known} 
tracker present: 53-72\% of random domains, 88-98\% of top websites, and 78-100\% of websites 
in the different Swedish top-categories.  While Google has by far the greatest third-party 
coverage among known tracker companies (e.g., Google has trackers on more than 90\% of the 
websites within the majority of the website categories considered here), we have found that 
there are many other third-party domains that may fly under the radar.  
For example, Disconnect's blocking list only detects 10\% of the 
%external 
third-party primary domains.  
With most of these non-blocked third-party entries being third-party content providers, which are 
known to keep track of users across services, we expect that there will be a continued 
battle for the knowledge about our web activities.

The remainder of the paper is organized as follows. 
%Section~\ref{sec:background} gives a brief background on online tracking.  
Our methodology and measurement framework are 
described in Section~\ref{sec:method}.  Section~\ref{sec:https} presents our 
characterization results for HTTPS adoption and client redirections between 
secure and non-secure resources.  Section~\ref{sec:tracking} characterizes the 
third-party tracking landscape and how it differs between secure and non-secure 
domains. Finally, Section~\ref{sec:related} discusses related works, before 
Section~\ref{sec:conclusions} concludes the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Background}\label{sec:background}

@@

%\begin{table*}[t]
%\centering
%\caption{Summary of domain lists.}
%\label{tab:domains}
%{\small
%\begin{tabular}{|l|c|r|l|r|r|}\hline
%\hline %% closing line
%Domain list       & Date       & Total size  & Selection & Selection size & Unique\\\hline
%.SE health status & 2014-03-27 & 980         & curated   & 915            & 915   \\\hline
%.se zone          & 2014-07-10 & 1,318,000   & random    & 100,000        & 100,000 \\\hline
%.dk zone          & 2014-07-23 & 1,260,000   & random    & 10,000         & 10,000 \\\hline
%.com zone         & 2014-08-27 & 114,178,000 & random    & 10,000         & 10,000 \\\hline
%.net zone         & 2014-08-27 & 15,096,000  & random    & 10,000         & 10,000 \\\hline
%reach50           & 2014-09-01 & 50          & top       & 50             & 50 \\\hline
%Alexa Top 1M      & 2014-09-01 & 1,000,000   & top       & 10,000         & 9,986 \\\hline
%                  &            &             & random    & 10,000         & 9,959 \\\hline
%                  &            &             & all .se   & 3,364          & 3 364 \\\hline
%                  &            &             & all .dk   & 2,637          & 2,637 \\\hline
%Total             &            & 132,852,050 &           & 156,907        & 156,045 \\\hline
%\end{tabular}}
%\vspace{-0pt}
%\end{table*}

@@

Our measurement campaign is as follows: 
we visit the front page of each domain, 
using both \texttt{http} and \texttt{https},
from a large set of domains.  
For each case we consider both the domain with and without the \texttt{www} prefix.  
Using the tool developed in this project, 
the front page of each domain is downloaded and parsed using the headless \texttt{phantomjs} browser 
the same way the users' browser would.
%\footnote{
%%Since the headless browser does not have a display window,
%As part of the validation process,
%the tool creates an image of the full website after the download is complete,
%but does not display the results to any user.}  
The URL of each requested resource is extracted, 
and then associated with the domain it was loaded from.  
Finally, each object and domain is classified and prepared 
for post processing and analysis.

%In this work, 
%We perform a large-scale measurement campaign in which 
%the front page of a large set of domains are visited
%using both \texttt{http} and \texttt{https}.
%For each cases,
%we also considered both domain variations 
%with and without the \texttt{www} prefix.
%With the tool developed in this project,
%%\footnote{The browser does not have a visible window, as it is built for automation.}
%%browser \texttt{phantomjs} downloads
%the front page of each domain is downloaded and parsed 
%using the headless 
%(without display window) 
%\texttt{phantomjs} browser 
%the same way a user's browser would.\footnote{
%Since the headless browser does not have a display window,
%As part of the validation process,
%the tool creates an image of the full website after the download is complete,
%but does not display the results to any user.}
%The URL of each requested resource is extracted, and associated with
%the domain it was loaded from. Each object and domain is classified 
%and prepared for post processing and analysis.

{\bf Swedish perspective:}  
First, our measurements are performed from computers located in Sweden.
Second, we leverage the domain lists from the .SE Health Status report\cite{LoWa12},
identified by the Internet Infrastructure Foundation (.SE) as the most important domains
to Swedish internet usage and operation.
Combined, these lists include approximately 1,000 domains in the
categories: counties (21), domain registrars (146), financial services (79), 
government-owned corporations (GOCS) (60), higher education (49), 
ISPs (20), media (33), municipalities (290), and public authorities (282).
Third, we consider the 50 global websites most visited by Swedish Internet users (reach50),
all .se domains (3,364) among the top-million globally most popular domains
according to Alexa\footnote{ Alexa (official website), \url{www.alexa.com}.}),
as well as 100K randomly selected websites within the .se domain zone.

{\bf Global baseline:}  
%To put findings in perspective and to broaden our
%conclusions we 
To put our findings into perspective and 
%as a result 
broaden our conclusions we
also consider the globally most popular websites (according to Alexa),
all .dk (Denmark) domains (2,637) among the top-million globally most popular domains,
%\footnote{Alexa (official website), \url{www.alexa.com}.}) 
and randomly selected websites from the .com, .net, and .dk (Denmark) domain zones.
Table~\ref{tab:domains} summarizes the domain lists used.
% and selection.
%%and any bias due to geographic personalization by content providers and
%%Content Delivery Networks (CDNs), for example.
%%Client in Sweden and section of websites ...

{\bf Data collection and parallelization:}
HTTP/HTTPS traffic metadata such
as requested URLs and their HTTP request/response headers have been recorded in the HTTP
Archive (HAR) data format.
During our data collection,
multiple domains have been retrieved in parallel, 
with parallelism adjusted to fit the available computer capacity. 
To reduce the risk of intermittent errors,
%% – either in software, on the network or in the remote system – 
each failed access has been retried up to two times.

\begin{table}[t]
\centering
\caption{Summary of domain lists.}
\label{tab:domains}
{\small
\begin{tabular}{|l|c|r|l|r|r|}
%%\cline{4-6}
\hline
\multicolumn{3}{|c}{Domain lists} & \multicolumn{3}{|c|}{Selection} \\\hline
List(s)           & Date       & Total size  & Type      & Size & Unique\\\hline
.SE health        & 27/3/14    & 980         & curated   & 915            & 915   \\\hline
.se zone          & 10/7/14    & 1,318,000   & random    & 100,000        & 100,000 \\\hline
.dk zone          & 23/7/14    & 1,260,000   & random    & 10,000         & 10,000 \\\hline
.com zone         & 27/8/14    & 114,178,000 & random    & 10,000         & 10,000 \\\hline
.net zone         & 27/8/14    & 15,096,000  & random    & 10,000         & 10,000 \\\hline
reach50           & 1/9/14     & 50          & top       & 50             & 50 \\\hline
Alexa 1M          & 1/9/14     & 1,000,000   & top       & 10,000         & 9,986 \\\hline
                  &            &             & random    & 10,000         & 9,959 \\\hline
                  &            &             & all .se   & 3,364          & 3 364 \\\hline
                  &            &             & all .dk   & 2,637          & 2,637 \\\hline
Total             &            & 132,852,050 &           & 156,907        & 156,045 \\\hline
\end{tabular}}
\vspace{-16pt}
\end{table}

@@

{\bf Data extraction:}
A custom-built tool based on the command line JSON processor \texttt{jq}
is used to convert the HAR data into JSON-based HAR data with information about both requests and responses.
The extracted data includes protocol and hosts (e.g., from URLs), 
HTTP status, mime-type, referer and redirect values – both
for the origin domain's front page and any resources requests by it.

{\bf Domain identification and classification:}
Each requested resource is classified across multiple dimensions, for example: 
it's mime type, if it was retrieved securely or not, and if it was downloaded 
from an internal sub-domain or from a known third-part tracker. 
%We classify each requested resource along many dimensions, 
%including its mime type, if it is retrieved securely,
%and if it is downloaded from an internal subdomain or from an known third-party tracker, for example. 
To determine the primary domain and subdomain(s) we use the 
public suffix list\footnote{Public suffix list, \url{https://publicsuffix.org/}, May 2015.} 
%(used by browser vendors to implement domain-dependent security measures)
and match domains against it.  This is important to capture the relationships
between second-level domains not open for commercial registration (e.g., \texttt{.com.br} under \texttt{.br}).  
We refer to resources requested from the main page
that are retrieved from the same domain, a subdomain
or a superdomain as {\em internal} requests.
Any other request is deemed as an {\em external} request.

{\bf Tracker identification:}
%When analyzing third-party tracking,
%it is important to note that 
Public lists of known trackers typically are incomplete and out-of-date~\cite{MPS+13, KrWi06}.
Potentially any external resource can be used for third-party tracking.
For example, 
Even static (non-script, non-executable) resources with no capabilities to
dynamically survey the users' browser or OS can track users across domains using
the referer HTTP header and customized URI's.
